---
title: "Career Development Tool for Young Professionals"
subtitle: "**Sample Recommendation Report (UserID==3)**"
output: pdf_document
papersize: a4
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_knit$set(root.dir = normalizePath("/Users/Shijie/Documents/A-Gatech/CS 6460 Education Foundation/homework/Project"))

library(ggplot2)
```


```{r echo=FALSE, results='hide',message=FALSE, warning=FALSE}
# Data Readin
job <- read.csv("original/google_job_skills.csv")
job <- job[ -which(job$Minimum.Qualifications==""), ]

# Create Job ID using row index
job$ID <- rownames(job)
n = ncol(job)
job <- job[ , c(n, 1:(n-1)) ]
```


```{r echo=FALSE, results='hide',message=FALSE, warning=FALSE}
# 1 Text Mining

# 1.1 Minimum Qualification
job_m <- job[, c(1,3,4,7,8)]

# 1.1.1 Required Degree
job_m$degree = NA

job_m$degree[ grepl("MBA", job_m$Minimum.Qualification)  ] = 4

job_m$degree[grepl("BA", job_m$Minimum.Qualification) & is.na(job_m$degree) ] = 1
job_m$degree[grepl("BS", job_m$Minimum.Qualification) & is.na(job_m$degree) ] = 1
job_m$degree[grepl("Bachelor", job_m$Minimum.Qualification) & is.na(job_m$degree) ] = 1
job_m$degree[ grepl("Business Intern", job_m$Title) & is.na(job_m$degree) ] = 1

job_m$degree[grepl("MA", job_m$Minimum.Qualification)] = 2
job_m$degree[grepl("MS", job_m$Minimum.Qualification)] = 2
job_m$degree[grepl("Master", job_m$Minimum.Qualification)] = 2

job_m$degree[grepl("PhD", job_m$Minimum.Qualification)] = 3
job_m$degree[grepl("Ph.D", job_m$Minimum.Qualification)] = 3

job_m$degree[grepl("JD", job_m$Minimum.Qualification)] = 5
job_m$degree[grepl("J.D.", job_m$Minimum.Qualification)] = 5


# 1.1.2 Required Years of Experience
job_m$Minimum.Qualifications <- gsub("\n", " ", job_m$Minimum.Qualifications)

pattern <- "([1-9]|[1-9][0-9])\\s(years|year|-year)\\sof"

job_m$expRaw <- substring( job_m$Minimum.Qualifications, regexpr(pattern, job_m$Minimum.Qualifications), regexpr(pattern, job_m$Minimum.Qualifications) + 2)

job_m$exp <- as.numeric( gsub("[A-z]", "", job_m$expRaw) )


# 1.2 Extracting Skills

job_m$q <-paste(job_m$Minimum.Qualifications, job_m$Preferred.Qualifications, sep = " ")

job_m$python = 0
job_m$python[grepl("Python", job_m$q)] = 1 

job_m$java = 0
job_m$java[grepl("Java", job_m$q)] = 1 

job_m$cpp = 0
job_m$cpp[grepl("C\\+\\+", job_m$q)] = 1 

job_m$c = 0
job_m$c[grepl("\\sC\\s", job_m$q)] = 1

job_m$chash = 0
job_m$chash[grepl("C#", job_m$q)] = 1

job_m$php = 0
job_m$php[grepl("PHP", job_m$q)] = 1 

job_m$javascript = 0
job_m$javascript[grepl("JavaScript", job_m$q)] = 1 

job_m$sql = 0
job_m$sql[grepl("SQL", job_m$q)] = 1

job_m$ruby = 0
job_m$ruby[grepl("Ruby", job_m$q)] = 1

job_m$perl = 0
job_m$perl[grepl("Perl", job_m$q)] = 1

# 1.3 Delete observations with all missing values in skills
job_m$t <- rowSums( job_m[, c( 10:ncol(job_m))], na.rm=TRUE)
job_m <- job_m[job_m$t>0,]

# 1.4 Create final matrix
job_m <- job_m[ , c( 1, 6, 8, 10: (ncol(job_m)-1) ) ]

```


```{r echo=FALSE, results='hide',message=FALSE, warning=FALSE}
# 2. Matching

# 2.1 Minimum requirement
filter.min.req <- function(i, data) {
  # i: an user
  # data: row: jobs and a target user; col: skills
  
  # degree
  d <- data[ data$ID==i, ]$degree
  if ( !is.na(d) ) { # has degree requirement
    data_f <- data[ !is.na(data$degree), ]
    data_f <- data_f[ data_f$degree>=d, ]
  }
 
  # Years of experience
  e <- data[ data$ID==i, ]$exp
  if ( !is.na(e) ) { # has work requirement
    data_f <- data_f[ !is.na(data_f$exp), ]
    data_f <- data_f[ data_f$exp>=e, ]
  }
  
  return(data_f)
}


# 2.2 KNN
k.nearest.neighbors <- function(i, data, k = 5, comp=FALSE ){  
  # i: an user
  # data: row: jobs and a target user; col: skills
  # k=5: offer 5 recommendations
  data <- t(data)
  similarities <- cor(data)
  distances <- -log((similarities / 2) + 0.5)
  var <- paste("X", i, sep="")
  distances<-data.frame(distances)
  
  list <- cbind(distances[, names(distances)==var], as.numeric(row.names(distances) ) ) 
  list <- list[order(list[,1]),]
  
  if(comp==FALSE) {
    return(list[,2] [2:(k + 1)])
  }else{
    return(distances)
  }
}


# 2.3 Job recommendation
user = 3

job_f <- filter.min.req(user, job_m)
job_f <- job_f[, 4:ncol(job_f)]
id <- k.nearest.neighbors(user, job_f, k=5, comp=FALSE)

id <-data.frame(id)
job_recommendation <- merge(id, job, by.x = "id", by.y = "ID")

# 2.4 Skill recommendation
skill_rec <- rbind(id, user)
skill_rec <- merge(skill_rec, job_m, by.x = "id", by.y = "ID")
skill_rec <- rbind(skill_rec, colMeans(skill_rec, na.rm = TRUE))

skill_comp <- rbind(skill_rec[skill_rec$id==as.character(user),], skill_rec[nrow(skill_rec),] )
skill_recommendation <- skill_comp[ , (skill_comp[2, ] > skill_comp[1,]) ]
skill_recommendation$exp <- round( as.numeric(skill_recommendation$exp), digits = 1)
skill_recommendation[2,1] <- ""
row.names(skill_recommendation) <- c("User", "Recommended")
```


```{r echo=FALSE, results='hide',message=FALSE, warning=FALSE}
# 3. Ranking - Competitiveness Analysis
# Scenario: an user (ID=3) wants to apply for a recommended job (ID = 24). What is his ranking?
user = 3
job = 24

var <- paste("X", job, sep="")

comp <- k.nearest.neighbors(user, job_f, k=5, comp=TRUE)
comp_list <- data.frame(comp[, names(comp)==var], as.numeric(row.names(comp) ) )

colnames(comp_list)<-c("score", "ID")
comp_list <- comp_list[order(comp_list[,1]),]
comp_list <- comp_list[-which(comp_list$ID==24), ]

# Histogram
pic <- ggplot(comp_list, aes(x=score)) + geom_histogram()

scoreLine <- comp_list[ comp_list[,comp_list$ID==user], 1]

comp_pic <- pic + geom_vline(aes(xintercept=scoreLine),
                 color="blue", linetype="dashed", size=1) + ggtitle("Competitiveness Analysis") 
```

## 1. Job recommendations

```{r message=FALSE}
job_recommendation[, 1:3]
```
The list above shows the top five jobs that most match the profile of the targeted user.

## 2. Skill recommendations
```{r message=FALSE}
skill_recommendation
```
The table above suggests that if the user wants to increase her competency for the targeted jobs, she may consider accumulating more years of work experience to advance her career further.  

## 3. Competitiveness analysis
<div style= "float:right;position: relative; top: -80px;">
```{r message=FALSE, fig.height=2.5, fig.width=3}
nrow(comp_list)
comp_pic
```
</div>
The picture above shows the user has an excellent profile that meets the requirement of the targeted job (job id equals to 24). However, about 20 other job seekers also have similar profiles as the user does. Considering there are 117 job seekers in the pool, the user's competency level for the targeted job is around the top 17