---
title: "auto_patent"
output: html_document
---

```{r}
rm(list=ls())
library(stringr)
library(Rcpp)

setwd("/Users/Shijie/Documents/A-Vandy/RA/Auto patent/Data")
load("patentsfull.rda")
df <- unique(patent.df.merged)
df <- df[, c(5,1,2,3,4)]

# sample data
# load("patentsample.rda")
# df <- unique(patent.sample)
```

# Split patent.sample$basic 
```{r}
# save the current column names to vector cur_colnames
cur_colnames=colnames(df)

# split patent$basic by \n, save the result into a character vector basic
basic<-unlist(strsplit(as.character(df$basic), "\n"))

# save the 1st, 11th, 21th...etc. elements of basic to the 6th column of patent.sample
# save the 2st, 12th, 22th...etc. elements of basic to the 7th column of patent.sample
# combine the 4st and 5th, 14th and 15th...etc. elements of basic, 
# then save to the 8th column of patent.sample
# save the 8st, 18th, 28th...etc. elements of basic to the 9th column of patent.sample

### seq
# for (i in 1:nrow(df)){
#   df[i,6] = basic[10*(i-1)+1]
#   df[i,7] = basic[10*(i-1)+2]
#   df[i,8] = paste(basic[10*(i-1)+4], basic[10*(i-1)+5])
#   df[i,9]= basic[10*(i-1)+8]
# } 
# 
# # rename the columns of patent.sample
# names(df) <- append(cur_colnames, c("basic_country", "basic_patent_num", "basic_inventor", "basic_date") )
```

# Define an index-based split function 
```{r}
# # define an index-based split function 
# indexSplit <- function(parentColumn, start, stop, end=FALSE ){
#   #check if end
#   if(end==FALSE){
#   split_result <- substr(parentColumn, start+max(0, attr(start, "match.length")), stop-1)
#   }else{
#    split_result <-substr(parentColumn, start+max(0, attr(start, "match.length")), stop)
#   }
#   return (split_result)
# }
```

# edge conditions: non-matching data
# define two index-redefining functions 
```{r}
# Step 1. redefine indexes
# check for non-matching data: -1 
# if not match, replace it with the next matching result
# e.g. observation 7, 92, 97
#   index_us_class index_cpc_class index_intl_class index_field_of_search index_c_end
# 7             1              28              194                    -1         310
# 92            1              47               -1                   127         181
# 97            1              67              118                    -1         232
# 
# we want to make the changes as below
#
#   index_us_class index_cpc_class index_intl_class index_field_of_search index_c_end
# 7             1              28              194                    310         310
# 92            1              47              127                    127         181
# 97            1              67              118                    232         232
indexReorg <- function(combined_index_df){
  for( i in (ncol(combined_index_df)-1) : 1 ){
    combined_index_df[,i] = ifelse( combined_index_df[,i]==-1,
                                    combined_index_df[,i+1],
                                    combined_index_df[,i])
    }
  return(combined_index_df)
}

# # Step 2. Mapping the redefined data back to original index matrixes
# indexMap<-function(combined_index_df, index_matrix, a){
#   for( i in 1:length(index_matrix) ){
#     index_matrix[i]=combined_index_df[i,a]
#   }
#   return (index_matrix)
# }
```



# Split patent.sample$inventor
```{r}
# find the indexes of "Inventors:", "Applicant:", "Assignee:", "Family ID:"
#             "Appl. No.:", "Filed:", and the end of the character string
index_inventor<-regexpr("Inventors:", df$inventor)
index_applicant<-regexpr("Applicant:", df$inventor)
index_assignee<-regexpr("Assignee:", df$inventor)
index_family_id<-regexpr("Family ID:", df$inventor)
index_appNo<-regexpr("Appl. No.:", df$inventor)
index_filed_date<-regexpr("Filed:", df$inventor)
index_i_end<-str_length(df$inventor) 
## index_i_end<-regexpr("([1-9]|[12]\\d|30|31), \\d{4}", patent.sample$inventor)

# construct the index matrix
index_matrix_i_df <- data.frame(index_inventor, index_applicant, 
                                 index_assignee, index_family_id, index_appNo,index_filed_date, index_i_end)
# edge conditions: non-matching data
# redefine the indexes
index_matrix_i_df= indexReorg(index_matrix_i_df)

# split patent$inventor
df$inventor_inventor = ifelse( index_matrix_i_df[,1] != index_matrix_i_df[,2], 
                                    substr(df$inventor, index_matrix_i_df[,1]+str_length("Inventors:"), index_matrix_i_df[,2]-1 ),
                                    substr(df$inventor, index_matrix_i_df[,1], index_matrix_i_df[,2])    )

df$inventor_applicant = ifelse( index_matrix_i_df[,2] != index_matrix_i_df[,3], 
                                    substr(df$inventor, index_matrix_i_df[,2]+str_length("Applicant:"), index_matrix_i_df[,3]-1 ),
                                    substr(df$inventor, index_matrix_i_df[,2], index_matrix_i_df[,3])    )

df$inventor_assignee = ifelse( index_matrix_i_df[,3] != index_matrix_i_df[,4], 
                                    substr(df$inventor, index_matrix_i_df[,3]+str_length("Assignee:"), index_matrix_i_df[,4]-1 ),
                                    substr(df$inventor, index_matrix_i_df[,3], index_matrix_i_df[,4])    )

df$inventor_family_id = ifelse( index_matrix_i_df[,4] != index_matrix_i_df[,5], 
                                    substr(df$inventor, index_matrix_i_df[,4]+str_length("Family ID:"), index_matrix_i_df[,5]-1 ),
                                    substr(df$inventor, index_matrix_i_df[,4], index_matrix_i_df[,5])    )

df$inventor_appNo = ifelse( index_matrix_i_df[,5] != index_matrix_i_df[,6], 
                                    substr(df$inventor, index_matrix_i_df[,5]+str_length("Appl. No.:"), index_matrix_i_df[,6]-1 ),
                                    substr(df$inventor, index_matrix_i_df[,5], index_matrix_i_df[,6])    )

df$inventor_filed_date = ifelse( index_matrix_i_df[,6] != index_matrix_i_df[,7], 
                                    substr(df$inventor, index_matrix_i_df[,6]+str_length("Filed:"), index_matrix_i_df[,7] ),
                                    substr(df$inventor, index_matrix_i_df[,6], index_matrix_i_df[,7])    )
  


# index_inventor = indexMap(index_matrix_i_df,  index_inventor, 1)
# index_applicant = indexMap(index_matrix_i_df,  index_applicant, 2)
# index_assignee = indexMap(index_matrix_i_df,  index_assignee, 3)
# index_family_id = indexMap(index_matrix_i_df, index_family_id, 4)
# index_appNo = indexMap(index_matrix_i_df, index_appNo, 5)
# index_filed_date = indexMap(index_matrix_i_df, index_filed_date, 6)
# 
# # split patent$inventor
# df$inventor_inventor = indexSplit(df$inventor, index_inventor, index_applicant)
# 
# df$inventor_applicant = indexSplit(df$inventor, index_applicant, index_assignee)
# df$inventor_assignee = indexSplit(df$inventor, index_assignee, index_family_id)
# df$inventor_family_id = indexSplit(df$inventor, index_family_id, index_appNo)
# df$inventor_appNo = indexSplit(df$inventor, index_appNo, index_filed_date)
# df$inventor_filed_date = indexSplit(df$inventor, index_filed_date, index_i_end, end=TRUE)
```


# Split patent.sample$class
```{r}
# find the indexes of "Current U.S. Class:", "Current CPC Class:", 
#   "Current International Class", "Field of Search:", and the end of the character string
index_us_class<-regexpr("Current U.S. Class:", df$class)
index_cpc_class<-regexpr("Current CPC Class:", df$class)
index_intl_class<-regexpr("Current International Class:", df$class)
index_field_of_search<-regexpr("Field of Search:", df$class)
index_c_end<-str_length(df$class)

# construct the index matrix
index_matrix_c_df <- data.frame(index_us_class, index_cpc_class, 
                                 index_intl_class, index_field_of_search, index_c_end)

# edge conditions: non-matching data
# redefine the indexes
index_matrix_c_df= indexReorg(index_matrix_c_df)


# split df$class
df$class_us_class = ifelse( index_matrix_c_df[,1] != index_matrix_c_df[,2], 
                                    substr(df$class, index_matrix_c_df[,1]+str_length("Current U.S. Class:"), index_matrix_c_df[,2]-1 ),
                                    substr(df$class, index_matrix_c_df[,1], index_matrix_c_df[,2])    )

df$class_cpc_class = ifelse( index_matrix_c_df[,2] != index_matrix_c_df[,3],
                                    substr(df$class, index_matrix_c_df[,2]+str_length("Current CPC Class:"), index_matrix_c_df[,3]-1 ),
                                    substr(df$class, index_matrix_c_df[,2], index_matrix_c_df[,3])    )

df$class_intl_class = ifelse( index_matrix_c_df[,3] != index_matrix_c_df[,4], 
                                    substr(df$class, index_matrix_c_df[,3]+str_length("Current International Class:"), index_matrix_c_df[,4]-1 ),
                                    substr(df$class, index_matrix_c_df[,3], index_matrix_c_df[,4])    )

df$class_field_of_search = ifelse( index_matrix_c_df[,4] != index_matrix_c_df[,5], 
                                    substr(df$class, index_matrix_c_df[,4]+str_length("Field of Search:"), index_matrix_c_df[,5] ),
                                    substr(df$class, index_matrix_c_df[,4], index_matrix_c_df[,5])    )

# index_us_class = indexMap(index_matrix_c_df,  index_us_class, 1)
# index_cpc_class = indexMap(index_matrix_c_df,  index_cpc_class, 2)
# index_intl_class = indexMap(index_matrix_c_df,  index_intl_class, 3)
# index_field_of_search = indexMap(index_matrix_c_df, index_field_of_search, 4)
# 
# df$class_us_class = indexSplit(df$class, index_us_class, index_cpc_class)
# df$class_cpc_class = indexSplit(df$class, index_cpc_class, index_intl_class)
# df$class_intl_class = indexSplit(df$class, index_intl_class, index_field_of_search)
# df$class_field_of_search = indexSplit(df$class, index_field_of_search, index_c_end, end=TRUE)

```


# remove whitespace and non-word characters from the whole date frame
```{r}

# remove the leading and trailing non-word characters: //W
 df$class_field_of_search = gsub("^\\W+|\\W+$", "", df$class_field_of_search)

# remove the leading and trailing whitespace: //s
# must perform this step before spliting 'reference' 
#     to avoid extra leading empty string after strsplit
for( i in 1:ncol(df) ) {
  df[, i] = gsub("^\\s+|\\s+$", "", df[, i])
}


```

# Split patent.sample$refernce
```{r}
# pre-condition: no leading and trailing whitespace in 'reference'
reference<-strsplit(as.character(df$reference), "\n")

# # examine the split result
# for (i in 1: length(reference)){
#   if ( !is_multiple3 (reference[[i]]) ){
#     print("Obeservation"+ df[i,1]+ "is not a multiple of 3.")
#   }
# }

```

# Split patent.sample$refernce
```{r}
# create a new data frame to save the splitted data from reference
patent_num<-df[,1]
data_size <-length(reference)
sourceCpp("/Users/Shijie/Documents/A-Vandy/RA/Auto patent/Rcode/patent_reference_split.cpp")
ref <- refSplit(reference, patent_num, data_size)
ref <- ref[ref[,1]!="", ]

# 
# cppFunction('
#     CharacterMatrix  refSplit( CharacterVector vector, String patentNum) {
#       int count=0;
#       int size = vector.size()/3;
#       CharacterMatrix out(size,4);
# 
#       for( int j=0; j<size; j++){
#         for(int k=0; k<3; k++){
#           out(j,0)=patentNum;
#           out(j,k+1)=vector[count];
#           count++;
#         }
#     }
#     return out;
#   }
# ')
# 
# 
# patent_num<-df[,1]
# re<-data.frame()
# for(i in 1:length(reference)){
#   patentNum<-patent_num[i]
#   ref_vector = reference[[i]]
#   ref_piece = refSplit(ref_vector, patentNum)
#   re = rbind(re, ref_piece)
# }


# patent_num<-df[,1]
# re<-data.frame()
# for(i in 1:length(reference) ){
#   ref_vector = reference[[i]]
#   count=1
#   for(j in 1:(length(ref_vector)/3)){
#     re_df<-c("", "", "", "")
#     for(k in 1:3){
#       re_df[1]=patent_num[i]
#       re_df[k+1]=ref_vector[count]
#       count=count+1
#     }
#     re = rbind(re, re_df)
#   }
# }


# ref_df<-data.frame()
# m=0
# for(i in 1:length(reference)){
#   ref_vector = reference[[i]]
#   count=1
#   for(j in 1:(length(ref_vector)/3)){
#     for(k in 1:3){
#       ref_df[j+m,1]=df[i,1]
#       ref_df[j+m,k+1]=ref_vector[count]
#       count=count+1
#     }
#   }
#   m=m+j
# }
# ref_df<- ref_df[order(ref_df$V1),]

# rename the columns of ref_df
names(ref) <- c("patent_num", "ref_num", "ref_date", "ref_inventor")

```

# Helper function
```{r}
# # check if the number of objects in a vector is the mutiple of 3
# is_multiple3 <- function(x){
#   if (x %% 3 == 0){
#     return TRUE
#   }
#   return FALSE
# }
```


# Save data
```{r}
save(df, file="/Users/Shijie/Documents/A-Vandy/RA/Auto patent/Data/patentfull_split.rda")
save(ref, file="/Users/Shijie/Documents/A-Vandy/RA/Auto patent/Data/patentfull_split_reference.rda")

```

# comment:
abnormal values: 1. check the cleaning algrithms. 2.If not, create an error list with patent number.
packag grep. regular expression.
try [:alnum:]  first
gsub grep, remove or extract, find the easier way
google chrome package:SelectGadget


