---
title: "auto_patent"
output: html_document
---

```{r}
rm(list=ls())
setwd("/Users/Shijie/Documents/A-Vandy/RA/Auto patent")
load("patentsample.rda")
# install.packages("stringr", dependencies = TRUE)
library(stringr)
df <- patent.sample
```

# Split patent.sample$basic 
```{r}
# save the current column names to vector cur_colnames
cur_colnames=colnames(df)

# split patent$basic by \n, save the result into a character vector basic
basic<-unlist(strsplit(as.character(df$basic), "\n"))

# save the 1st, 11th, 21th...etc. elements of basic to the 6th column of patent.sample
# save the 2st, 12th, 22th...etc. elements of basic to the 7th column of patent.sample
# combine the 4st and 5th, 14th and 15th...etc. elements of basic, 
# then save to the 8th column of patent.sample
# save the 8st, 18th, 28th...etc. elements of basic to the 9th column of patent.sample

### seq
for (i in 1:nrow(df)){
  df[i,6] = basic[10*(i-1)+1]
  df[i,7] = basic[10*(i-1)+2]
  df[i,8] = paste(basic[10*(i-1)+4], basic[10*(i-1)+5])
  df[i,9]= basic[10*(i-1)+8]
} 

# rename the columns of patent.sample
names(df) <- append(cur_colnames, c("basic_country", "basic_patent_num", "basic_inventor", "basic_date") )
```

# Define an index-based split function 
```{r}
# define an index-based split function 
indexSplit <- function(parentColumn, start, stop, end=FALSE ){
  #check if end
  if(!end){
  split_result <- substr(parentColumn, start+max(0, attr(start, "match.length")), stop-1)
  }else{
   split_result <-substr(parentColumn, start+max(0, attr(start, "match.length")), stop)
  }
  return (split_result)
}
```

# edge conditions: non-matching data
# define two index-redefining functions 
```{r}
# Step 1. redefine indexes
# check for non-matching data: -1 
# if not match, replace it with the next matching result
# e.g. observation 7, 92, 97
#   index_us_class index_cpc_class index_intl_class index_field_of_search index_c_end
# 7             1              28              194                    -1         310
# 92            1              47               -1                   127         181
# 97            1              67              118                    -1         232
# 
# we want to make the changes as below
#
#   index_us_class index_cpc_class index_intl_class index_field_of_search index_c_end
# 7             1              28              194                    310         310
# 92            1              47              127                    127         181
# 97            1              67              118                    232         232
indexReorg <- function(combined_index_df){
  for( i in 1:nrow(combined_index_df) ){
    for( j in 1:ncol(combined_index_df) ){
      while( combined_index_df[i,j]==-1 ){
        combined_index_df[i,j]=combined_index_df[i,j+1]
      }
    }
  }
  return(combined_index_df)
}

# Step 2. Mapping the redefined data back to original index matrixes
indexMap<-function(combined_index_df, index_matrix, a){
  for( i in 1:length(index_matrix) ){
    index_matrix[i]=combined_index_df[i,a]
  }
  return (index_matrix)
}
```



# Split patent.sample$inventor
```{r}
# find the indexes of "Inventors:", "Applicant:", "Assignee:", "Family ID:"
#             "Appl. No.:", "Filed:", and the end of the character string
index_inventor<-regexpr("Inventors:", df$inventor)
index_applicant<-regexpr("Applicant:", df$inventor)
index_assignee<-regexpr("Assignee:", df$inventor)
index_family_id<-regexpr("Family ID:", df$inventor)
index_appNo<-regexpr("Appl. No.:", df$inventor)
index_filed_date<-regexpr("Filed:", df$inventor)
index_i_end<-str_length(df$inventor) ## index_i_end<-regexpr("([1-9]|[12]\\d|30|31), \\d{4}", patent.sample$inventor)

# construct the index matrix
index_matrix_i_df <- data.frame(index_inventor, index_applicant, 
                                 index_assignee, index_family_id, index_appNo,index_filed_date, index_i_end)
# edge conditions: non-matching data
# redefine the indexes
index_matrix_i_df= indexReorg(index_matrix_i_df)
index_inventor = indexMap(index_matrix_i_df,  index_inventor, 1)
index_applicant = indexMap(index_matrix_i_df,  index_applicant, 2)
index_assignee = indexMap(index_matrix_i_df,  index_assignee, 3)
index_family_id = indexMap(index_matrix_i_df, index_family_id, 4)
index_appNo = indexMap(index_matrix_i_df, index_appNo, 5)
index_filed_date = indexMap(index_matrix_i_df, index_filed_date, 6)

# split patent$inventor
df$inventor_inventor = indexSplit(df$inventor, index_inventor, index_applicant)
df$inventor_applicant = indexSplit(df$inventor, index_applicant, index_assignee)
df$inventor_assignee = indexSplit(df$inventor, index_assignee, index_family_id)
df$inventor_family_id = indexSplit(df$inventor, index_family_id, index_appNo)
df$inventor_appNo = indexSplit(df$inventor, index_appNo, index_filed_date)
df$inventor_filed_date = indexSplit(df$inventor, index_filed_date, index_i_end, end=TRUE)
```


# Split patent.sample$class
```{r}
# find the indexes of "Current U.S. Class:", "Current CPC Class:", 
#   "Current International Class", "Field of Search:", and the end of the character string
index_us_class<-regexpr("Current U.S. Class:", df$class)
index_cpc_class<-regexpr("Current CPC Class:", df$class)
index_intl_class<-regexpr("Current International Class:", df$class)
index_field_of_search<-regexpr("Field of Search:", df$class)
index_c_end<-str_length(df$class)

# construct the index matrix
index_matrix_c_df <- data.frame(index_us_class, index_cpc_class, 
                                 index_intl_class, index_field_of_search, index_c_end)

# edge conditions: non-matching data
# redefine the indexes
index_matrix_c_df= indexReorg(index_matrix_c_df)
index_us_class = indexMap(index_matrix_c_df,  index_us_class, 1)
index_cpc_class = indexMap(index_matrix_c_df,  index_cpc_class, 2)
index_intl_class = indexMap(index_matrix_c_df,  index_intl_class, 3)
index_field_of_search = indexMap(index_matrix_c_df, index_field_of_search, 4)

df$class_us_class = indexSplit(df$class, index_us_class, index_cpc_class)
df$class_cpc_class = indexSplit(df$class, index_cpc_class, index_intl_class)
df$class_intl_class = indexSplit(df$class, index_intl_class, index_field_of_search)
df$class_field_of_search = indexSplit(df$class, index_field_of_search, index_c_end, end=TRUE)

```


# remove whitespace and non-word characters from the whole date frame
```{r}
# remove the leading and trailing whitespace: //s
# must perform this step before spliting 'reference' 
#     to avoid extra leading empty string after strsplit
for( i in 1:ncol(df) ) {
  df[, i] = gsub("^\\s+|\\s+$", "", df[, i])
}

# remove the leading and trailing non-word characters: //W
 df$class_field_of_search = gsub("^\\W+|\\W+$", "", df$class_field_of_search)
```

# Split patent.sample$refernce
```{r}
# pre-condition: no leading and trailing whitespace in 'reference'
reference<-strsplit(as.character(df$reference), "\n")

# # examine the split result
# for (i in 1: length(reference)){
#   if ( !is_multiple3 (reference[[i]]) ){
#     print("Obeservation"+ df[i,1]+ "is not a multiple of 3.")
#   }
# }

```

# Split patent.sample$refernce
```{r}
# create a new data frame to save the splitted data from reference
ref_df<-data.frame()
m=0
for(i in 1:length(reference) ){
  ref_vector = reference[[i]]
  count=1
  for(j in 1:(length(ref_vector)/3)){
    for(k in 1:3){
      ref_df[j+m,1]=df[i,1]
      ref_df[j+m,k+1]=ref_vector[count]
      count=count+1
    }
  }
  m=m+j
}

# rename the columns of ref_df
names(ref_df) <- c("patent_num", "ref_num", "ref_date", "ref_inventor")

```

# Helper function
```{r}
# # check if the number of objects in a vector is the mutiple of 3
# is_multiple3 <- function(x){
#   if (x %% 3 == 0){
#     return TRUE
#   }
#   return FALSE
# }
```


# Save data
```{r}
write.csv(df, "patentsample_split.csv")
write.csv(ref_df, "patentsample_reference_split.csv")
```

# comment:
abnormal values: 1. check the cleaning algrithms. 2.If not, create an error list with patent number.
packag grep. regular expression.
try [:alnum:]  first
gsub grep, remove or extract, find the easier way
google chrome package:SelectGadget


