---
title: 'Project 2: Modeling and Evaluation'
subtitle: |-
  CSE6242 - Data and Visual Analytics - Fall 2017
  Due: Sunday, November 26, 2017 at 11:59 PM UTC-12:00 on T-Square
output:
  pdf_document: default
  html_notebook:
    code_folding: none
    theme: default
  html_document:
    code_folding: none
    theme: default
---

__Name: Shijie Shi __

__GaTech ID: SSHI48 __

__Email: shijie.shi@gatech.edu __

# Data

We will use the same dataset as Project 1: [`movies_merged`](https://s3.amazonaws.com/content.udacity-data.com/courses/gt-cs6242/project/movies_merged).

# Objective

Your goal in this project is to build a linear regression model that can predict the `Gross` revenue earned by a movie based on other variables. You may use R packages to fit and evaluate a regression model (no need to implement regression yourself). Please stick to linear regression, however.

# Instructions

You should be familiar with using an [RMarkdown](http://rmarkdown.rstudio.com) Notebook by now. Remember that you have to open it in RStudio, and you can run code chunks by pressing *Cmd+Shift+Enter*.

Please complete the tasks below and submit this R Markdown file (as **pr2.Rmd**) containing all completed code chunks and written responses, and a PDF export of it (as **pr2.pdf**) which should include the outputs and plots as well.

_Note that **Setup** and **Data Preprocessing** steps do not carry any points, however, they need to be completed as instructed in order to get meaningful results._

# Setup

Same as Project 1, load the dataset into memory:

```{r}
load('movies_merged')
```

This creates an object of the same name (`movies_merged`). For convenience, you can copy it to `df` and start using it:

```{r}
df = movies_merged
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
colnames(df)
```

## Load R packages

Load any R packages that you will need to use. You can come back to this chunk, edit it and re-run to load any additional packages later.

```{r}
library(ggplot2)
library(stringr)
library(GGally)
library(tm)
library(lubridate)
```

If you are using any non-standard packages (ones that have not been discussed in class or explicitly allowed for this project), please mention them below. Include any special instructions if they cannot be installed using the regular `install.packages('<pkg name>')` command.

**Non-standard packages used**: None

# Data Preprocessing

Before we start building models, we should clean up the dataset and perform any preprocessing steps that may be necessary. Some of these steps can be copied in from your Project 1 solution. It may be helpful to print the dimensions of the resulting dataframe at each step.

## 1. Remove non-movie rows

```{r}
# TODO: Remove all rows from df that do not correspond to movies
df = subset(df, Type=='movie')
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
```

## 2. Drop rows with missing `Gross` value

Since our goal is to model `Gross` revenue against other variables, rows that have missing `Gross` values are not useful to us.

```{r}
# TODO: Remove rows with missing Gross value
df = subset(df, !is.na(Gross))
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
```

## 3. Exclude movies released prior to 2000

Inflation and other global financial factors may affect the revenue earned by movies during certain periods of time. Taking that into account is out of scope for this project, so let's exclude all movies that were released prior to the year 2000 (you may use `Released`, `Date` or `Year` for this purpose).

```{r}
# TODO: Exclude movies released prior to 2000
df = subset(df, Year >= 2000)
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
```

## 4. Eliminate mismatched rows

_Note: You may compare the `Released` column (string representation of release date) with either `Year` or `Date` (numeric representation of the year) to find mismatches. The goal is to avoid removing more than 10% of the rows._

```{r}
# TODO: Remove mismatched rows
release_year = as.numeric(substr(df$Released,1,4))
# df_test = subset(df, Year==release_year)
# cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
# We will remove more than 631 rows if we remove all mismatched Year and release_year.

# I choose to remove Year and released_year differ only by 1 
df = subset(df, abs(Year-release_year)<=1)
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
# 116 rows removed, less than 10% of the original rows (3332)
```

## 5. Drop `Domestic_Gross` column

`Domestic_Gross` is basically the amount of revenue a movie earned within the US. Understandably, it is very highly correlated with `Gross` and is in fact equal to it for movies that were not released globally. Hence, it should be removed for modeling purposes.

```{r}
# TODO: Exclude the `Domestic_Gross` column
df <- subset(df,	select	= -Domestic_Gross)
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
```

## 6. Process `Runtime` column

```{r}
# TODO: Replace df$Runtime with a numeric column containing the runtime in minutes

# Test if data in df$Runtime are all saved as minutes
time_unit = substr(df$Runtime, str_length(df$Runtime)-3, str_length(df$Runtime))
unique(time_unit)
# [1] " min" "N/A" 

# In my data, the time saved in df$Runtime are all minutes (no hour, or other time units)
df$runtime= as.numeric( substr(df$Runtime, 1, str_length(df$Runtime)-3) ) 
```

Perform any additional preprocessing steps that you find necessary, such as dealing with missing values or highly correlated columns (feel free to add more code chunks, markdown blocks and plots here as necessary).

```{r}
# TODO(optional): Additional preprocessing

# delete the data where Gross is missing or non-positive
df <- subset(df, Gross>0)
```

_**Note**: Do NOT convert categorical variables (like `Genre`) into binary columns yet. You will do that later as part of a model improvement task._

## Final preprocessed dataset

Report the dimensions of the preprocessed dataset you will be using for modeling and evaluation, and print all the final column names. (Again, `Domestic_Gross` should not be in this list!)

```{r}
# TODO: Print the dimensions of the final preprocessed dataset and column names
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
colnames(df)
```

# Evaluation Strategy

In each of the tasks described in the next section, you will build a regression model. In order to compare their performance, you will compute the training and test Root Mean Squared Error (RMSE) at different training set sizes.

First, randomly sample 10-20% of the preprocessed dataset and keep that aside as the **test set**. Do not use these rows for training! The remainder of the preprocessed dataset is your **training data**.

Now use the following evaluation procedure for each model:

- Choose a suitable sequence of training set sizes, e.g. 10%, 20%, 30%, ..., 100% (10-20 different sizes should suffice). For each size, sample that many inputs from the training data, train your model, and compute the resulting training and test RMSE.
- Repeat your training and evaluation at least 10 times at each training set size, and average the RMSE results for stability.
- Generate a graph of the averaged train and test RMSE values as a function of the train set size (%), with optional error bars.

You can define a helper function that applies this procedure to a given set of features and reuse it.

# Tasks

Each of the following tasks is worth 20 points, for a total of 100 points for this project. Remember to build each model as specified, evaluate it using the strategy outlined above, and plot the training and test errors by training set size (%).

## 1. Numeric variables

Use Linear Regression to predict `Gross` based on available _numeric_ variables. You can choose to include all or a subset of them.

```{r}
# TODO: Build & evaluate model 1 (numeric variables only)
df_1 <- df[complete.cases(df$imdbRating), ]

model_1 = "Gross~Budget+imdbRating"

aveRMSE	<- function(percen, df, model){
		ave_sample_train=c()
		ave_sample_test=c()
		
		for(i	in	1:10){
		sample_size	<- floor(percen	* nrow(df))
		set.seed(1)
		
		#split	data	into	train	and	test	data
		spl	<- sample(1:nrow(df),	size	= sample_size)
		sample_train	<- df[spl,	]
		sample_test	<- df[-spl,	]
		
		#building	linear	regression	model	based	on train	data
		Model_1=lm(model, sample_train)
		
		#obtain	test set predictions	from	Model_1
		yhat=predict(Model_1,	sample_test)
		yhat=as.data.frame(yhat)
		RMSE_test=sqrt(mean((sample_test$Gross-yhat)^2))
		
		
		
		ave_sample_train=c(ave_sample_train,sqrt(residuals(Model_1)^2))
		ave_sample_test=c(ave_sample_test,RMSE_test)
		}
		final=c()
		final=c(final,mean(ave_sample_train))
		final=c(final,mean(ave_sample_test))
		return (final)
}

RMSE_vector <- function(perc_vector, df, model, train){
  if (train==TRUE){
    trainRMSE=c();
    for( i in 1: length(perc_vector)) {
      trainRMSE= c( trainRMSE, aveRMSE( perc_vector[i], df, model )[1] )
    }
    return (trainRMSE)
  }else {
    testRMSE=c();
    for( i in 1: length(perc_vector)) {
      testRMSE= c( testRMSE, aveRMSE( perc_vector[i], df, model )[2] )
    }
    return(testRMSE)
  }
}


percentage <- seq(0.1,0.9,by=0.1)
trainRMSE <- RMSE_vector(percentage, df_1, model_1, train = TRUE)
testRMSE <- RMSE_vector(percentage, df_1, model_1, train = FALSE)

#Generate	a	graph	of	the	averaged	train	and	test	RMSE	as	a	function	of the	train	set	size	(%)
result_1<- data.frame(percentage,	trainRMSE,testRMSE)
ggplot(result_1,	aes(percentage))	+
		geom_line(aes(y	= trainRMSE,	color="train"))	+
		geom_line(aes(y	= testRMSE,	color="test")) +
    scale_x_continuous(breaks=seq(0.1,0.9,0.1), labels=scales::percent) +
    ylab("RMSE") +
    xlab("Training Set Sizes")

# Report the minimum test RMSE
min(testRMSE)
    
```

**Q**: List the numeric variables you used.

**A**: The numeric variables I used are: "Budget" and "imdbRating".

**Q**: What is the best mean test RMSE value you observed, and at what training set size?

**A**: The best mean test RMSE is 105498947 This value is observed when the training set is 80% of the original dataset.


## 2. Feature transformations

Try to improve the prediction quality from **Task 1** as much as possible by adding feature transformations of the numeric variables. Explore both numeric transformations such as power transforms and non-numeric transformations of the numeric variables like binning (e.g. `is_budget_greater_than_3M`).

```{r}
# TODO: Build & evaluate model 2 (transformed numeric variables only)

ggplot(df_1,	aes(x=Budget,	y=Gross))	+
				geom_point(shape=1)	+				
				geom_smooth()

ggplot(df_1,	aes(x=Budget^1.1,	y=Gross))	+
				geom_point(shape=1)	+				
				geom_smooth()

ggplot(df_1,	aes(x=imdbRating,	y=Gross))	+
				geom_point(shape=1)	+				
				geom_smooth()
ggplot(df_1,	aes(x=imdbRating^5,	y=Gross))	+
				geom_point(shape=1)	+				
				geom_smooth()



df_2<-df_1
df_2$Budget <- df_1$Budget^1.1
df_2$imdbRating <- df_1$imdbRating^5
model_2 = "Gross~Budget+imdbRating"

# df_2$imdbRating <- df_1$imdbRating>5.0
# test RMSE didn't drop when binning "imdbRating"

trainRMSE <- RMSE_vector(percentage, df_2, model_2, train = TRUE)
testRMSE <- RMSE_vector(percentage, df_2, model_2, train = FALSE)


#Generate	a	graph	of	the	averaged	train	and	test	RMSE	as	a	function	of the	train	set	size	(%)
result_2<- data.frame(percentage,	trainRMSE,testRMSE)
ggplot(result_2,	aes(percentage))	+
		geom_line(aes(y	= trainRMSE,	color="train"))	+
		geom_line(aes(y	= testRMSE,	color="test")) +
    scale_x_continuous(breaks=seq(0.1,0.9,0.1), labels=scales::percent) +
    ylab("RMSE") +
    xlab("Training Set Sizes")

# Report the minimum test RMSE
min(testRMSE)

```

**Q**: Explain which transformations you used and why you chose them.

**A**: I did power transformation for both Budget and imdbRatings. Budget was transformed to Budget^1.1. imdbRatings was transformed to imdbRatings^5. For Budget, I tried to change the power in a sequence of 1 to 2 by 0.1 then find 1.1 is the optimal power that produces the lowerst test RMSE. For imdbRatings, I tried to change the power in a sequence of 1 to 6 by 0.5 then find 5 is the optimal power that produces the lowerst test RMSE. 

I have considered to do binning for imdbRating (imdbRatings is greater than 5 or not) but the test RMSE increased after the bining. 


**Q**: How did the RMSE change compared to Task 1?

**A**: The RMSE dropped from 105498947 to 104733372.


## 3. Non-numeric variables

Write code that converts genre, actors, directors, and other categorical variables to columns that can be used for regression (e.g. binary columns as you did in Project 1). Also process variables such as awards into more useful columns (again, like you did in Project 1). Now use these converted columns only to build your next model.

```{r}
# TODO: Build & evaluate model 3 (converted non-numeric variables only)

# 3.1 Replace Genre with a collection of binary columns
df_3 <- df_2
df_3 <- df_3[complete.cases(df_3$Genre), ]

data <- Corpus(VectorSource(df_3$Genre))
dataTM<- DocumentTermMatrix( tm_map(data, removePunctuation)  )

binaryMatrix <- function(x) {
  r <- nrow(x)
  c <- ncol(x)
  y <- matrix(vector(typeof(x$v), 1 * r * c), r, c)
  y[cbind(x$i, x$j)] <- x$v
  dimnames(y) <- x$dimnames
  return(y)
}

genre_binary <- as.data.frame( binaryMatrix(dataTM) )
df_3 <- cbind(df_3, genre_binary)
top10 <- data.frame(sort(colSums(df_3[,40:62]),decreasing = TRUE)[1:10])
top10



# 3.2 Convert Awards to 2 numeric columns: wins and nominations
countWin <- function(x){
  if(x!= "n/a"){
    count = 0
    if(str_detect(x,"win")){
      a <- as.numeric(gregexpr("win",x))
      b <- substring(x, a-3, a)
      c <- gregexpr('[0-9]+', b)
      count = count + as.numeric(regmatches(b,c)[[1]])
    }
    if(str_detect(x,"won")){
      a <- as.numeric(gregexpr("won",x))
      b <- substring(x, a+3, a+6)
      c <- gregexpr('[0-9]+', b)
      count = count + as.numeric(regmatches(b,c)[[1]])
    }
    return(count)
  }
  else(return(NA))
}

countNomi<- function(x){
  if(x!= "n/a"){
    n_nom = 0
    if(str_detect(x,"nomination")){
      a <- as.numeric(gregexpr("nomination",x))
      b <- substring(x, a-4, a)
      c <- gregexpr('[0-9]+', b)
      n_nom = n_nom + as.numeric(regmatches(b,c)[[1]])
    }
    if(str_detect(x,"nominated for")){
      a <- as.numeric(gregexpr("nominated for",x))
      b <- substring(x, a+13, a+16)
      c <- gregexpr('[0-9]+', b)
      n_nom = n_nom + as.numeric(regmatches(b,c)[[1]])
    }
    return(n_nom)
  }
  else(return(NA))
}

award <- df_3$Awards
award <- sapply(award, tolower)
df_3$Win <- sapply(award, countWin)
df_3$Nomination <- sapply(award, countNomi)
summary(df_3$Win>0)
summary(df_3$Nomination>0)
# Replace all NAs in Win and Nomination with 0. (Here I assumn that movies has missing data in Awards can be counted as movies that did not win any monimations and awards. )
df_3$Win[is.na(df_3$Win)] = 0
df_3$Nomination[is.na(df_3$Nomination)] = 0

cor(df_3$Win, df_3$Nomination)
# [1] 0.7978272




model_3 = "Gross~drama+comedy+action+adventure+romance+crime+thriller+horror+mystery+fantasy+Win"

# df_2$imdbRating <- df_1$imdbRating>5.0
# test RMSE didn't drop when binning "imdbRating"

trainRMSE <- RMSE_vector(percentage, df_3, model_3, train = TRUE)
testRMSE <- RMSE_vector(percentage, df_3, model_3, train = FALSE)


#Generate	a	graph	of	the	averaged	train	and	test	RMSE	as	a	function	of the	train	set	size	(%)
result_3<- data.frame(percentage,	trainRMSE,testRMSE)
ggplot(result_3,	aes(percentage))	+
		geom_line(aes(y	= trainRMSE,	color="train"))	+
		geom_line(aes(y	= testRMSE,	color="test")) +
    scale_x_continuous(breaks=seq(0.1,0.9,0.1), labels=scales::percent) +
    ylab("RMSE") +
    xlab("Training Set Sizes")

min(testRMSE)
```

**Q**: Explain which categorical variables you used, and how you encoded them into features.

**A**: The categorical variables I used are: drama, comedy, action, adventure, romance, crime, thriller, horror, mystery, fantasy, and Win. These variables represents movies from top 10 most common genres and the total number of awards awarded to the movie. The genre types are encoded into binary variables where 0 represents the movie does not belong to this type and 1 represents the movie belongs to this type.

I find high coorelation between the total number of awards received and the total number of nominations so I only have the total number awards received (df_3$Win) added to my model. 


**Q**: What is the best mean test RMSE value you observed, and at what training set size? How does this compare with Task 2?

**A**: The best mean test RMSE is 136961078. This result is received when the training set contains 90% of the original data. The mean test RMSE increased by 32227706 (it was 104733372 in step 2). 


## 4. Numeric and categorical variables

Try to improve the prediction quality as much as possible by using both numeric and non-numeric variables from **Tasks 2 & 3**.

```{r}
# TODO: Build & evaluate model 4 (numeric & converted non-numeric variables)

model_4 = "Gross~Budget+imdbRating+drama+comedy+action+adventure+romance+crime+thriller+horror+mystery+fantasy+Win"

trainRMSE <- RMSE_vector(percentage, df_3, model_4, train = TRUE)
testRMSE <- RMSE_vector(percentage, df_3, model_4, train = FALSE)


#Generate	a	graph	of	the	averaged	train	and	test	RMSE	as	a	function	of the	train	set	size	(%)
result_4<- data.frame(percentage,	trainRMSE,testRMSE)
ggplot(result_4,	aes(percentage))	+
		geom_line(aes(y	= trainRMSE,	color="train"))	+
		geom_line(aes(y	= testRMSE,	color="test")) +
    scale_x_continuous(breaks=seq(0.1,0.9,0.1), labels=scales::percent) +
    ylab("RMSE") +
    xlab("Training Set Sizes")

min(testRMSE)
```

**Q**: Compare the observed RMSE with Tasks 2 & 3.

**A**: The observed RMSE is 100636018. Compare to Task 2, the observed RMSE dropped by 4097354.  Compare to Task 3, the ovserved RMSE dropped by 36325060.



## 5. Additional features

Now try creating additional features such as interactions (e.g. `is_genre_comedy` x `is_budget_greater_than_3M`) or deeper analysis of complex variables (e.g. text analysis of full-text columns like `Plot`).

```{r}
# TODO: Build & evaluate model 5 (numeric, non-numeric and additional features)

df_5 <-df_3
df_5 $is_budget_greater_than_3M = 1*(df_5$Budget>3000000)

model_5 = "Gross~Budget+imdbRating+drama+comedy+action+adventure+romance+crime+thriller+horror+mystery+fantasy+Win + Budget*fantasy"
# 99656822

# model_5 = "Gross~Budget+imdbRating+drama+comedy+action+adventure+romance+crime+thriller+horror+mystery+fantasy+Win + is_budget_greater_than_3M*fantasy"
# 100641863

trainRMSE <- RMSE_vector(percentage, df_5, model_5, train = TRUE)
testRMSE <- RMSE_vector(percentage, df_5, model_5, train = FALSE)

#Generate	a	graph	of	the	averaged	train	and	test	RMSE	as	a	function	of the	train	set	size	(%)
result_5<- data.frame(percentage,	trainRMSE,testRMSE)
ggplot(result_5,	aes(percentage))	+
		geom_line(aes(y	= trainRMSE,	color="train"))	+
		geom_line(aes(y	= testRMSE,	color="test")) +
    scale_x_continuous(breaks=seq(0.1,0.9,0.1), labels=scales::percent) +
    ylab("RMSE") +
    xlab("Training Set Sizes")

min(testRMSE)
```

**Q**: Explain what new features you designed and why you chose them.

**A**: 	I	used interactions between	the `Budget` and `fantasy`. I tried to encode `Budget` as binary varaibles such as `is_budget_greater_than_3M` but this change did not improve the model. The intersaction between `Budget` and `fantasy` represent the additional gross that is associated with fantasy movies.  


**Q**: Comment on the final RMSE values you obtained, and what you learned through the course of this project.

**A**: I am not very happy with the final RMSE I got. There are still a lot of options to improve the model. This projects helps me to practice how to develop regression models in data analysis progress.

